---
title: "260-2017-04-24-speed-lab"
author: Rick Gilmore
date: "`r Sys.time()`"
bibliography: bib/bibliography.bib
output:
  ioslides_presentation:
    css: css/ioslides.css
    incremental: no
    transition: default
    widescreen: yes
  pdf_document:
    latex_engine: xelatex
  word_document: default
---

## Today's topic

- Measuring the speed of nervous system conduction
- And a tiny lesson in open, transparent, reproducible data science

## Question

- How fast does the nervous system conduct information?
- [Prior evidence](https://psu-psychology.github.io/psych-260-spring-2017/lecture-notes/260-2017-03-29-somatosensation.html#15)
    - Proprioception vs. touch

## Prediction

We predict that average speed of conduction will be ...
    
## Scheme

- Speed = Distance/Time
- Chain of participants to make distance larger
    + If typical person ~ 1.5 m, then 
    + at s=30 m/s, t = d/s -> 1.5/30 = `r 1.5/30` secs.

## Condition 1 (ankle)

- Squeeze ankle
- ankle_shoulder + shoulder_brain + brain_decide + brain_shoulder + shoulder_hand

## Condition 2 (shoulder)

- Squeeze shoulder
- shoulder_brain + brain_decide + brain_shoulder + shoulder_hand
- Condition 1 - Condition 2
- ankle_shoulder + ~~shoulder_brain~~ + ~~brain_decide~~ + ~~brain_shoulder~~ + ~~shoulder_hand~~

## Measure

- sum(ankle_shoulder) for all participants -> Distance
- mean(time(Condition 1)) - mean(time(Condition 2)) -> Time
- Speed = Distance/Time

## Materials

- Stop watch
- Tape measure

## Decisions

- Same hand or dominant?
- Alternate ankle/shoulder or one condition before the other?
- How many trials?
    + Fixed number?
    + When reach asymptote?
    
## Data files

- Data file with [body measurements](https://docs.google.com/spreadsheets/d/1NFGu-M4AGf_4IHf-o4y3sCHUZ3Ao-CpBUpXqCsQrAIg/edit#gid=0)
    + participant, ankleshoulder (cm)
- Data file with [reaction times](https://docs.google.com/spreadsheets/d/1NFGu-M4AGf_4IHf-o4y3sCHUZ3Ao-CpBUpXqCsQrAIg/edit#gid=1626241513)
    + trial {1...n}, condition {ankle, shoulder}, time (s)

```{r set-up-r, echo=FALSE}
library("googlesheets")
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("ggplot2"))
```

## Measuring distance

```{r load-distance}
psych260 <- gs_title("psych-260-spring-2017")
psych260 %>% 
  gs_read(ws = "distance") ->
  distance
dist.hist <- ggplot(data = distance, aes(x=ankleshoulder)) +
  geom_histogram(bins = 5)
```

----

```{r measuring-distance, echo = FALSE}
dist.hist
```

## Sum distance

```{r}
with(distance, summary(ankleshoulder))

# Calculate sum
dist.sum = with(distance, sum(ankleshoulder))
```

The total distance is `r dist.sum` cm.

## Measuring time

```{r load-times}
psych260 %>% 
  gs_read(ws = "time") ->
  time

# Plot data
time.plot = ggplot(data = time, aes(x=trial, y=time, color=condition)) +
  geom_point() + 
  geom_line()
```


----

```{r measuring-time, echo = FALSE}
time.plot
```

## Calculate time difference

```{r calc-time-diffs}
time %>% 
  filter(condition == "ankle") -> 
  ankle.times

time %>% 
  filter(condition == "shoulder") -> 
  shoulder.times

time.diff <- data_frame(trial=unique(time$trial),
                        time=ankle.times$time - shoulder.times$time)

time.diff.plot = ggplot(data = time.diff, aes(x=trial, y=time)) +
  geom_point() + 
  geom_line()
```

----

```{r condition-diff, echo = FALSE}
time.diff.plot
```

## Calculating speed

```{r}
time.diff$speed <- dist.sum/time.diff$time

speed.hist <- ggplot(data = time.diff, aes(x=speed)) +
  geom_histogram(bins = 5)
```

---

```{r speed-summary, echo = FALSE}
speed.hist
```

## Plot time series of speeds

```{r plot-speed}
speed.plot <- ggplot(data = time.diff, aes(x=trial, y=speed)) +
  geom_point() +
  geom_line()
```

---

```{r}
speed.plot
```

## Summarizing findings

- We tested the mean speed of neural propagation in a sample of n=`r dim(distance)[1]` college-age adults.
- The mean speed of neural propagation over `r max(time.diff$trial)` trials was `r with(time.diff, summary(speed))['Mean']/100` m/s with a range of [`r with(time.diff, summary(speed))['Min.']/100`, `r with(time.diff, summary(speed))['Max.']/100`] m/s.
- These findings are/are not generally in accord with values we would expect from the literature.

## Limitations

## How to replicate/extend

## Resources {.smaller}

This document was prepared in RStudio 1.0.36 on `r Sys.time()`.

```{r session-info}
sessionInfo()
```

